---
title: "B. rapa annotation"
output: html_notebook
---

Want to get Arabidopsis best hits for each B.rapa gene.

Use the UCD B.rapa CDS fasta file,  [File S4 Devistty et al (G3)](http://www.g3journal.org/content/4/11/2065.supplemental)

BLAST against [TAIR10 CDS](ftp://ftp.arabidopsis.org/home/tair/Sequences/blast_datasets/TAIR10_blastsets/TAIR10_cds_20110103_representative_gene_model_updated)

Because these organisms are relatively closely related I will use blastn instead of blastp to focus on best match

Get sequences
```{bash}
wget -nv ftp://ftp.arabidopsis.org/home/tair/Sequences/blast_datasets/TAIR10_blastsets/TAIR10_cds_20101214_updated
wget -nv http://www.g3journal.org/highwire/filestream/472465/field_highwire_adjunct_files/6/FileS4.zip
```

```{bash}
mv TAIR10_cds_20101214_updated At.TAIR10.cds.fa
unzip -o FileS4.zip
makeblastdb -in At.TAIR10.cds.fa -dbtype nucl
```

```{bash}
blastn -query Brassica_rapa_final_CDS.fa \
  -db At.TAIR10.cds.fa \
  -strand both \
  -task dc-megablast \
  -outfmt 10 \
  -culling_limit 1 \
  -max_target_seqs 1 \
  -max_hsps 1 \
  -evalue 10e-4 \
  -num_threads 3 \
  -template_type coding \
  -template_length 16 \
  -out Brapa_vs_At_dc-megablast_out.csv
head Brapa_vs_At_dc-megablast_out.csv
```

Compare hits vs queries
```{bash}
grep -c ">" Brassica_rapa_final_CDS.fa # total number of query sequences
wc -l Brapa_vs_At_dc-megablast_out.csv # total number of queries with a hit
```

So 39709 of 44239 queries had a hit.  Seems reasonable

Import results
```{r}
brapa.blast <- read.csv("Brapa_vs_At_dc-megablast_out.csv", header = FALSE,as.is = TRUE)
colnames(brapa.blast) <- c("query","subject","perc_ID","aln_length","mismatch","gap_open","qstart","qend","sstart","send","eval","score")
head(brapa.blast)
summary(brapa.blast)
```

Add annotations from TAIR

Install database (only needs to be done once)
```{r, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("org.At.tair.db")
```


```{r}
library(org.At.tair.db)
brapa.blast$AGI <- sub("\\..*$","",brapa.blast$subject)

brapa.blast$At_symbol <- mapIds(org.At.tair.db,
       keys=brapa.blast$AGI,
       column = "SYMBOL",
       keytype = "TAIR",
       multiVals = function(x) paste(x,collapse = ";") )

brapa.blast$At_description <- 
  mapIds(org.At.tair.db,
       keys=brapa.blast$AGI,
       column = "GENENAME",
       keytype = "TAIR",
       multiVals = function(x) paste(x,collapse = ";") )

head(brapa.blast)
```

Now add location information

convert big bed to bed
```{bash}
bigBedtoBed brassica_rapa_v15_final_transcripts.bb brassica_rapa_v15_final_transcripts.bed
```


```{r}
Br.bed <- read.delim("brassica_rapa_v15_final_transcripts.bed",header=FALSE,as.is = TRUE)
colnames(Br.bed) <- unlist(strsplit("chrom start end name score strand thickStart thickEnd itemRgb blockCount blockSizes blockStarts",split=" "))
head(Br.bed)
summary(Br.bed)
```

```{r}
Brapa.annotated <- merge(Br.bed[,1:4],brapa.blast,by.x="name",by.y="query",all.x=TRUE)
Brapa.annotated <- Brapa.annotated[,unlist(strsplit("name chrom start end subject AGI At_symbol At_description perc_ID aln_length mismatch gap_open qstart qend sstart send eval score",split = " "))] # reorder columns 
Brapa.annotated <- Brapa.annotated[order(Brapa.annotated$chrom, Brapa.annotated$start),] 
head(Brapa.annotated)
```

```{r}
library(boxr)
box_auth()
write.csv(Brapa.annotated,"Brapa_V1.5_annotated.csv")
box_write(Brapa.annotated,filename = "Brapa_V1.5_annotated.csv.gz", dir_id = 24503987860)
```

