---
title: "G x E expression"
output: html_notebook
---

The question is how much G x E we see in the RNAseq data.  

To get started on this I ran the file "ProcessCount.R".  For this the data was TMM normalized, voom transformed, and then a gtXenvironment model was fit in limma.  

```{r}
library(limma)
library(tidyverse)
library(stringr)
library(modelr)
library(broom)
library(multidplyr)
library(VCA)
```

Get the limma objects
```{r}
load("voom.int.Rdata")
```

How many genes show GxE?

generate vector of interaction coefficients and then topTable
```{r, eval=FALSE}
colnames(fit.int$coefficients)
int.coef.n <- colnames(fit.int$coefficients) %>% str_which(":") 
topTable(fit.int,coef = int.coef.n)
```

Not working

let's just do an ANOVA for each gene

```{r}
counts <- as.tibble(counts.voom.gr$E) %>% 
  mutate(geneIndex=1:n())
counts[1:10,1:10]
```

```{r}
counts.long <- counts %>% gather(key="name",value="expression", -geneIndex) %>%
  mutate(gt=factor(str_extract(name,"RIL_[0-9]*")), 
         env=factor(str_extract(name,"CR|UN")), 
         rep = factor(str_extract(name,"Rep[0-9]"))) %>%
  select(-name) %>%
  group_by(geneIndex) %>% nest()
counts.long$data[[1]] 
#test.df <- counts.long$data[[1]] 
```

Now fit an anova and do a VCA for each gene in parallel
```{r, eval=FALSE}
fitAnova <- function(df) aov(expression ~ gt*env,data=df)

getPvals <- function(fit) {
  pvals <- summary(fit)[[1]]$`Pr(>F)`
  names(pvals) <- dimnames(summary(fit)[[1]])[[1]] %>% str_trim()
  data.frame(t(pvals))
}

fitVCA <- function(df) {
  tmpvca <- anovaVCA(expression ~ gt*env,Data=as.data.frame(df))
  data.frame(t(tmpvca$aov.tab[,"%Total"]))
}

cl <- create_cluster(cores=4)
counts.long <- partition(counts.long,geneIndex,cluster = cl)
cluster_library(cl,c("purrr","broom","VCA","stringr"))
cluster_copy(cl,fitAnova)
cluster_copy(cl,getPvals)
cluster_copy(cl,fitVCA)

system.time(
  counts.results <- counts.long %>% mutate(aov = map(data,fitAnova), 
                                        glance = map(aov,glance),
                                        pvals = map(aov,getPvals),
                                        varcomp = map(data,fitVCA)) %>%
    select(-data, -aov)
)
system.time(cluster_rm(cl,"counts.long"))
rm(counts.long)

```

```{r, eval=FALSE}
system.time(counts.results <- collect(counts.results))
save(counts.results,file="GxE_models.Rdata")
```

```{r}
if (! exists("counts.results")) load("GxE_models.Rdata")
aovsummary <- counts.results %>% unnest(glance,.drop = TRUE)
pvals <- counts.results %>% unnest(pvals,.drop = TRUE)
varcomp <- counts.results %>% unnest(varcomp,.drop = TRUE)
head(aovsummary)
head(pvals)
head(varcomp)
```

## PVals
Number of genes (out of 24619) that have a FDR adjusted pvalue of less than 0.05:
```{r}
pvals %>% ungroup() %>%
  summarise(gt=sum(p.adjust(gt,method="fdr") < 0.05),
            env=sum(p.adjust(env) < 0.05),
            gt.env=sum(p.adjust(gt.env,method="fdr") < 0.05))
```

## Variance components

### Histograms of % variance explained

```{r}
pl <- varcomp %>%
  ungroup() %>%
  select(-total) %>%
  gather(key = "component", value = "percent.variance",-geneIndex) %>%
  ggplot(aes(x=percent.variance)) +
  facet_wrap(~ component, scale="free_y") +
  geom_histogram()
pl
```

focus in on gt x env histogram, non-zeros
```{r}
pl <- varcomp %>%
  ungroup %>%
  filter(gt.env > 0) %>%
  ggplot(aes(x=gt.env)) +
  geom_histogram(binwidth = 5)
pl
```

Number of genes with gt.env explaining X % variance:
```{r}
varcomp$gt.env %>%
  cut(breaks=seq(0,35,5),right = FALSE) %>% table
```

