---
title: "Cr and UN Mutual Rank Clustering"
output: html_notebook
---

Notebook for generating mutual rank (MR) clusters from the CR and UN data separately.

```{r}
library(boxr)
library(magrittr)
library(knitr)
library(limma)
library(snowfall)
library(igraph)
library(plyr)
library(dplyr)
box_auth()
box_setwd(24503987860)
box_ls()
```

### First attempt, use RIL means
Get data
```{r}
#box_load(162362592785)
load("~/Box Sync/BrapaNetworks/Voom_CR_UN.Rdata")
annotation  <- read.csv("~/Box Sync/BrapaNetworks/Brapa_V1.5_annotated.csv.gz")
```


```{r}
RIL.coefs.CR <- grep("RIL",colnames(coef(voom.fit.CR)))
RIL.pvals.CR <- topTable(voom.fit.CR,coef = RIL.coefs.CR,number = Inf)
sum(RIL.pvals.CR$adj.P.Val<0.01)
sum(RIL.pvals.CR$adj.P.Val<10e-05) 
```

```{r}
RIL.coefs.UN <- grep("RIL",colnames(coef(voom.fit.UN)))
RIL.pvals.UN <- topTable(voom.fit.UN,coef = RIL.coefs.UN,number = Inf)
sum(RIL.pvals.UN$adj.P.Val<0.01)
sum(RIL.pvals.UN$adj.P.Val<10e-05) 
```
So lots of genes are highly significant, especially in UN

Start by taking the top 10000

```{r}
genes.of.interest.CR <- row.names(RIL.pvals.CR)[1:10000]
expr.data.CR <- voom.fit.CR$coefficients[genes.of.interest.CR,]
dim(expr.data.CR)
head(expr.data.CR[,1:6])

genes.of.interest.UN <- row.names(RIL.pvals.UN)[1:10000]
expr.data.UN <- voom.fit.UN$coefficients[genes.of.interest.UN,]
dim(expr.data.UN)
head(expr.data.UN[,1:6])
```

How many of these genes are in common in the two sets?

```{r}
sum(genes.of.interest.CR %in% genes.of.interest.UN)
```

```{r}
expr.data.CR[,-1] <- expr.data.CR[,-1] + expr.data.CR[,1] # add the intercept to the coefficients
colnames(expr.data.CR) <- sub("pdata.CR$RIL","",colnames(expr.data.CR),fixed=TRUE)
head(expr.data.CR[,1:6])

expr.data.UN[,-1] <- expr.data.UN[,-1] + expr.data.UN[,1] # add the intercept to the coefficients
colnames(expr.data.UN) <- sub("pdata.UN$RIL","",colnames(expr.data.UN),fixed=TRUE)
head(expr.data.UN[,1:6])
```

Get Rob's Data
```{r}
blups2011 <- read.csv("~/Box Sync/BrapaNetworks/Brapa2011BayesHeight_blups.csv",row.names = 1)
blups2012 <- read.csv("~/Box Sync/BrapaNetworks/Brapa2012BayesHeight_blups.csv", row.names = 1)
blups2011$Line %<>% paste("RIL",.,sep="_")
blups2012$Line %<>% paste("RIL",.,sep="_")
blups2011 %<>% subset(select=!grepl("V1|individual|blk|trt|treat",colnames(.)))
blups2012 %<>% subset(select=!grepl("V1|individual|blk|trt|treat",colnames(.)))
head(blups2011)
head(blups2012)
```

Have decided to use 2012 only...
```{r}
#blups.all <- merge(blups2011,blups2012,by="Line",suffixes = c(".2011",".2012"))
#head(blups.all)
blups.CR <- blups2012 %>% subset(select=grep("Line|CR",colnames(.)))
head(blups.CR)
blups.UN <- blups2012 %>% subset(select=grep("Line|UN",colnames(.)))
head(blups.UN)
```


Merge blups with expression
```{r}
expr.blup.CR <- merge(blups.CR,t(expr.data.CR),by.x="Line",by.y=0)
expr.blup.UN <- merge(blups.UN,t(expr.data.UN),by.x="Line",by.y=0)
head(expr.blup.CR[,1:6])
head(expr.blup.UN[,1:6])
```

Correlation
```{r}
expr.blup.cor.CR <- cor(expr.blup.CR[,-1])
expr.blup.cor.UN <- cor(expr.blup.UN[,-1])
```

Mutual Rank
```{r}
sfInit(parallel = TRUE, cpus=4)

expr.blup.rank.CR <- sfApply(expr.blup.cor.CR,2,function(x) rank(-abs(x)))
expr.blup.rank.CR[1:10,1:10]
expr.blup.mr.CR <- sqrt(expr.blup.rank.CR*t(expr.blup.rank.CR))
expr.blup.mr.CR[1:10,1:10]

expr.blup.rank.UN <- sfApply(expr.blup.cor.UN,2,function(x) rank(-abs(x)))
expr.blup.rank.UN[1:10,1:10]
expr.blup.mr.UN <- sqrt(expr.blup.rank.UN*t(expr.blup.rank.UN))
expr.blup.mr.UN[1:10,1:10]

sfStop()
```


```{r}
save(blups.CR,blups.UN,expr.blup.CR, expr.blup.UN, expr.blup.mr.CR, expr.blup.mr.UN,expr.data.CR,expr.data.UN,file = "~/Box Sync/BrapaNetworks/MR_CR_UN.Rdata")
```

Get graphs at different MR

## get data

```{r}
load("~/Box Sync/BrapaNetworks/MR_CR_UN.Rdata")
annotation <- read.csv("Brapa_V1.5_annotated.csv",row.names=1,as.is = TRUE)
```

## define some functions

```{r mr_subgraphs, echo=FALSE, cache.lazy=FALSE, warning=FALSE, eval=TRUE }
get.mr.subgraph <- function(mr.cutoff,mr.matrix,annotation=NULL,neighborhood,order=1) {
  #function to extract LFY graph at a specificed correlation cutoff
  gene.mr.tmp <- mr.matrix
  gene.mr.tmp[abs(gene.mr.tmp) > mr.cutoff] <- 0
  gene.mr.tmp[is.na(gene.mr.tmp)] <- 0 #important! otherwise vertices with NA edges are connected
  
  gene.graph <- graph.adjacency(adjmatrix = gene.mr.tmp,
                                mode="undirected",
                                weighted="mr",
                                diag=FALSE)
  
  #clust.membership <- clusters(gene.graph)$membership
  
  #colbar <- rainbow(max(clust.membership)+1)                 #define colors
  #V(gene.graph)$color <- colbar[clust.membership+1]             #assign colors to nodes
  sub.graphs <- graph.neighborhood(gene.graph,order=order,nodes=neighborhood) #for each ndoe of interest get all other nodes within order of 1
  
  #get combined list of vertices...
  
  sub.vertices <- unique(names(unlist(lapply(sub.graphs, V))))
  
  combined.sub.graph <- induced_subgraph(gene.graph,sub.vertices)
  
  V(combined.sub.graph)$color <- "lightblue"
  V(combined.sub.graph)[neighborhood]$color <- "red"
  if(!is.null(annotation)) V(combined.sub.graph)$gene <- 
    annotation$At_symbol[match(V(combined.sub.graph)$name,annotation$name)]
  list(cutoff=mr.cutoff,graph=combined.sub.graph,nodes=V(combined.sub.graph)$name)
}
```

## Get blup networks at different MR thresholds
Use different MR thresholds of 10 to 210 (absolute values) to build the network (genes with a correlation higher than the threshold were considered connected).  I then subset the network to take all genes "directly" connected to LFY.

```{r lfy_sub, echo=FALSE, eval=TRUE}
cutoffs <- c(10,20,30,50,80)
blup.mr.graphs.CR <- lapply(cutoffs, get.mr.subgraph, mr.matrix = expr.blup.mr.CR, annotation= annotation, neighborhood = colnames(blups.CR)[-1])
names(blup.mr.graphs.CR) <- sapply(blup.mr.graphs.CR, function(x) paste("blup.mr.graph.CR",x$cutoff,sep="."))

blup.mr.graphs.UN <- lapply(cutoffs, get.mr.subgraph, mr.matrix = expr.blup.mr.UN, annotation= annotation, neighborhood = colnames(blups.UN)[-1])
names(blup.mr.graphs.UN) <- sapply(blup.mr.graphs.UN, function(x) paste("blup.mr.graph.UN",x$cutoff,sep="."))
```

## plots

```{r plot, echo=FALSE, cache=TRUE}
cutoff=200
pdf("~/Box Sync/BrapaNetworks/blup.mr.plots.CR.pdf",height=12,width=12)
result <- sapply(blup.mr.graphs.CR,function (blup.graph) {
  if(blup.graph$cutoff <= cutoff) {
    E(blup.graph$graph)$weight <- rank(E(blup.graph$graph)$mr)^(1/4)
    plot(blup.graph$graph,
         layout = layout_with_kk, 
         vertex.label = ifelse(is.na(V(blup.graph$graph)$gene)|V(blup.graph$graph)$gene=="",
                               V(blup.graph$graph)$name,
                          paste(V(blup.graph$graph)$gene,V(blup.graph$graph)$name,sep="\n")),
         vertex.label.cex=1,
         main=paste("MR cutoff =",sub("blup.mr.graph.","",blup.graph$cutoff,fixed=T)))
    rm(blup.graph)
  }
})
dev.off()

cutoff=200
pdf("~/Box Sync/BrapaNetworks/blup.mr.plots.UN.pdf",height=12,width=12)
result <- sapply(blup.mr.graphs.UN,function (blup.graph) {
  if(blup.graph$cutoff <= cutoff) {
    E(blup.graph$graph)$weight <- rank(E(blup.graph$graph)$mr)^(1/4)
    plot(blup.graph$graph,
         layout = layout_with_kk, 
         vertex.label = ifelse(is.na(V(blup.graph$graph)$gene)|V(blup.graph$graph)$gene=="",
                               V(blup.graph$graph)$name,
                          paste(V(blup.graph$graph)$gene,V(blup.graph$graph)$name,sep="\n")),
         vertex.label.cex=1,
         main=paste("MR cutoff =",sub("blup.mr.graph.","",blup.graph$cutoff,fixed=T)))
    rm(blup.graph)
  }
})
dev.off()
```

make annotation tables for plots

```{r}
CR.annotation <- lapply(blup.mr.graphs.CR, function (blup.graph) {
    if(blup.graph$cutoff <= cutoff) {
      data.frame(
        MR_Cutoff=blup.graph$cutoff,
        trt="CR",
        annotation[annotation$name %in% blup.graph$nodes,]
      )
    }
}
)
CR.annotation <- ldply(CR.annotation)

write.csv(CR.annotation,file="~/Box Sync/BrapaNetworks/MR_CR_graphs_annotation.csv")

UN.annotation <- lapply(blup.mr.graphs.UN, function (blup.graph) {
    if(blup.graph$cutoff <= cutoff) {
      data.frame(
        MR_Cutoff=blup.graph$cutoff,
        trt="UN",
        annotation[annotation$name %in% blup.graph$nodes,]
      )
    }
}
)

UN.annotation <- ldply(UN.annotation)

write.csv(UN.annotation,file="~/Box Sync/BrapaNetworks/MR_UN_graphs_annotation.csv")

```

## Permutations

What is a reasonable threshold?  To find out, permute blups relative to gene expression, make graphs at different thresholds, and return number of genes in network

```{r}
permute.graphs <- function(blups,expression,cutoffs=cutoffs,permute=TRUE) {
  if(permute) blups$Line <- sample(blups$Line,replace = FALSE) # permute!
  expr.blup <- merge(blups,t(expression),by.x="Line",by.y=0)
  expr.blup.mr <- expr.blup[,-1] %>% 
    cor %>%  
    apply(2,function(x) rank(-abs(x))) %>%
    multiply_by(t(.)) %>%
    sqrt()
  blup.mr.graphs <- lapply(cutoffs, get.mr.subgraph, mr.matrix = expr.blup.mr, neighborhood = colnames(blups)[-1])
  sapply(blup.mr.graphs,function(graph) sum(grepl("^Bra",graph$nodes)))
}

test <- permute.graphs(blups.CR,expr.data.CR,cutoffs=c(10,20,30,50),permute=FALSE)
test
# should be the same as the first 4 of
sapply(blup.mr.graphs.CR,function(graph) sum(grepl("^Bra",graph$nodes)))

```


```{r, eval=TRUE}
permutations <- 100

sfInit(parallel=TRUE,cpus = 4)

sfLibrary(igraph)
sfLibrary(magrittr)
sfExport("blups.CR","blups.UN","expr.data.CR","expr.data.UN","permute.graphs","get.mr.subgraph","cutoffs")
system.time(perm.results.CR.100 <- sfSapply(1:permutations, function(p) {
  permute.graphs(blups.CR,expr.data.CR,cutoffs=cutoffs)
}))

save(perm.results.CR.100, file="permutation_results.Rdata")

perm.results.UN.100 <- sfSapply(1:permutations, function(p) {
  permute.graphs(blups.UN,expr.data.UN,cutoffs=cutoffs)
})

sfStop()

save(perm.results.UN.100, perm.results.CR.100, file="permutation_results.Rdata")
```

```{r}
load("permutation_results.Rdata")
apply(perm.results.CR.100,1,quantile,.95)
apply(perm.results.UN.100,1,quantile,.95)
```

```{r}
kable(cbind(threshold=cutoffs,
            "0.95 expectation"=apply(perm.results.CR.100,1,quantile,.95),
            observed=sapply(blup.mr.graphs.CR,function(graph) sum(grepl("^Bra",graph$nodes)))
), caption="Permutation analysis: CR network",digits=0)
```
```{r}
kable(cbind(threshold=cutoffs,
            "0.95 expectation"=apply(perm.results.UN.100,1,quantile,.95),
            observed=sapply(blup.mr.graphs.UN,function(graph) sum(grepl("^Bra",graph$nodes)))
), caption="Permutation analysis: UN network",digits=0)
```

Our networks are significant, but not sure that this is not helping set a threshold.  Try correlation with scale-free?
