---
title: "WGCNA"
output: html_notebook
---

WGCNA analysis of sun and shade expression in B. rapa RILs.

```{r}
library(tidyverse)
library(WGCNA)
library(ggplot2)
library(edgeR)
library(magrittr)
library(gplots)
options(stringsAsFactors = FALSE)
```

Get data
```{r}
#box_load(162362592785)
load("~/Box Sync/BrapaNetworks/Voom_CR_UN.Rdata")
annotation  <- read.csv("~/Box Sync/BrapaNetworks/Brapa_V1.5_annotated.csv.gz")
```

```{r}
RIL.coefs.UN <- grep("RIL",colnames(coef(voom.fit.UN)))
RIL.pvals.UN <- topTable(voom.fit.UN,coef = RIL.coefs.UN,number = Inf)
sum(RIL.pvals.UN$adj.P.Val<0.01)
sum(RIL.pvals.UN$adj.P.Val<10e-05) 
```
So lots of genes are highly significant, especially in UN

Start by taking the top 10000

```{r}
genes.of.interest.UN <- row.names(RIL.pvals.UN)[1:10000]
expr.data.UN <- voom.fit.UN$coefficients[genes.of.interest.UN,]
dim(expr.data.UN)
head(expr.data.UN[,1:6])
```


```{r}
expr.data.UN[,-1] <- expr.data.UN[,-1] + expr.data.UN[,1] # add the intercept to the coefficients
colnames(expr.data.UN) <- sub("pdata.UN$RIL","",colnames(expr.data.UN),fixed=TRUE)
colnames(expr.data.UN)[1] <- "RIL_1"
head(expr.data.UN[,1:6])
```

Get Rob's Data
```{r}
blups2012 <- read.csv("~/Box Sync/BrapaNetworks/Brapa2012BayesHeight_blups.csv", row.names = 1)
blups2012$Line %<>% paste("RIL",.,sep="_")
blups2012 %<>% subset(select=!grepl("V1|individual|blk|trt|treat",colnames(.)))
head(blups2012)
```

```{r}
blups.UN <- blups2012 %>% select(starts_with("Line"),starts_with("UN"),-matches("Inflection_size"))
head(blups.UN)
```

## WGCNA 

### UN

transform
```{r}
head(expr.data.UN[,1:6])
expr.data.UN.t <- t(expr.data.UN)
head(expr.data.UN.t[,1:6])
```

check sample quality

```{r}
gag.UN <- goodSamplesGenes(expr.data.UN.t, verbose = 3)
gag.UN$allOK
```

cluster samples to look for outliers

```{r,fig.width=12}
sampleTREE.UN <- hclust(dist(expr.data.UN.t), method = "average")
plot(sampleTREE.UN,cex=.6)
heatmap.2(expr.data.UN.t,Rowv=as.dendrogram(sampleTREE.UN), scale="col", trace="none")
```
So a few samples are different but they don't seem crazy different; leave them in for now.

Soft thresholding
```{r}
powers <- c(c(1:10), seq(from = 12, to=20, by=2))
sft <- pickSoftThreshold(expr.data.UN.t, powerVector = powers, verbose = 5)
```

```{r}
sizeGrWindow(9, 5)
par(mfrow = c(1,2))
cex1 <- 0.9
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

choose 5

```{r}
softPower <- 5
adjacency <- adjacency(expr.data.UN.t, power = softPower)
# Turn adjacency into topological overlap
TOM <- TOMsimilarity(adjacency);
dissTOM <- 1-TOM
```

```{r}
# Call the hierarchical clustering function
geneTree <- hclust(as.dist(dissTOM), method = "average")
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04)
```

define modules

```{r}
# We like large modules, so we set the minimum module size relatively high:
minModuleSize <- 30;
# Module identification using dynamic tree cut:
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit <- 2, pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize);
table(dynamicMods)
```

```{r}
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
```

merge similar modules

```{r}
# Calculate eigengenes
MEList <- moduleEigengenes(expr.data.UN.t, colors = dynamicColors)
MEs <- MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs);
# Cluster module eigengenes
METree <- hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
```

merge with correlation > 0.8
```{r}
MEDissThres = 0.2
# Plot the cut line into the dendrogram
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
abline(h=MEDissThres, col = "red")
# Call an automatic merging function
merge = mergeCloseModules(expr.data.UN.t, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs
```

compare pre and post merge
```{r}
sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
#dev.off()
```

```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs
```

```{r}
head(MEs)
dim(MEs)
dim(blups.UN)
rownames(expr.data.UN.t)
rownames(MEs)
rownames(MEs) <- rownames(expr.data.UN.t)
head(blups.UN)
```

```{r}
ME.blups <- merge(blups.UN,MEs,by.x=1.,by.y=0)
head(ME.blups)
```

```{r}
blups.cor <- cor(ME.blups[,2:6],ME.blups[,7:66])
max(blups.cor)
blups.cor
```


```{r}
blups.cor.P <- corPvalueStudent(blups.cor,nrow(expr.data.UN.t))
blups.cor.sig <- blups.cor
blups.cor.sig[blups.cor.P>0.05] <- NA
blups.cor.sig
```

plot it
```{r,fig.width=12}
# Will display correlations and their p-values
par(mar = c(8, 5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = blups.cor,
               yLabels = names(blups.UN),
               xLabels = names(MEs),
               xSymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               #textMatrix = signif(blups.cor.P,2),
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("UN Module-trait relationships"))

pdf("Module-trait_heatmap_UN.pdf",height = 8,width = 12)
par(mar = c(8, 5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = blups.cor,
               yLabels = names(blups.UN),
               xLabels = names(MEs),
               xSymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
              # textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("UN Module-trait relationships"))
dev.off()
```

Arbitrary, but let's take the max and min for each trait (so long as they are significant)
```{r}
blups.cor.5 <- blups.cor.sig

cor.top <- t(apply(blups.cor.5,1,function(x) {
  maxx = max(x,na.rm=TRUE)
  minx = min(x,na.rm=TRUE)
  ifelse(x == maxx | x == minx, x, NA)
}
))


cor.top <- cor.top[,apply(cor.top,2,function(x) !all(is.na(x)))]
cor.top
write.csv(cor.top,"Eigen_trait_cor_UN_.5_threshold.csv")
```

write the Eigen genes
```{r}
head(MEs[,colnames(cor.top)])
write.csv(MEs[,colnames(cor.top)],"Top_Eigen_genes_UN.csv")
```

