---
title: "BEL1 SNPs"
output: html_notebook
---

Are there any SNPs between IMB211 and R500 in BEL1?  Asking for a reviewer....

Bel1 is Bra025497 A04:8699173-8703567

BAMfiles are at /iplant/home/shared/ucd.plantbio/maloof.lab/members/upendra/RILs_mapping

Subset the BAM files:

    for f in $(ls *bam)
        do
        echo $f
        samtools index $f
        samtools view -b $f "A04:8699173-8703567" > subset/$f
        samtools index subset/$f
        done

Run Freebayes

    freebayes  -f  ~/Sequences/ref_genomes/B_rapa/genome/V1.5/Brapa_sequence_v1.5.fa mapping_v1.5_merged_bam_files/*.bam > bel1.vcf 
    
Run Snpeff

    java -jar snpEff/snpEff.jar  Brassica_rapa bel1.vcf > bel1_ann.vcf
    
Bring it into R

```{r}
library(tidyverse)
```

    
```{r}
vcf.data <- read_tsv("../output/bel1_ann.vcf",na = c("","NA","."),comment="#",col_names = FALSE)
```

```{r}
vcf.header <- system("grep '#C' ../output/bel1_ann.vcf",intern = TRUE) #might not work on Windows
vcf.header
vcf.header <- vcf.header %>% str_replace("#","") #get rid of the pound sign
vcf.header <- vcf.header %>% str_split(pattern = "\t")
colnames(vcf.data) <- vcf.header[[1]] #we need the [[1]] because str_split returns a list and we want the first element
View(vcf.data)
```

get the ANN column:

